Pomace: A Grappa for Non-Volatile Memory \vspace{-.2in}

### 2012-12-12 doc/nvm2013/paper.tex
mods to conclusion

\input{header}
\usepackage{microtype}
\usepackage{placeins}
\begin{document}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{.7in}
\addtolength{\oddsidemargin}{-.25in}
\addtolength{\evensidemargin}{-.25in}
\addtolength{\textwidth}{.5in}
\linespread{.94}

@@

\FloatBarrier
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{../asplos13/figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} Grappa system overview}
\end{minipage}
\vspace{-3ex}
\end{center}
\end{figure}

@@

\paragraph{\textls*{\small SSD}s assume locality.} In our benchmarks, we find the average request size to be
less than 64 bytes. Today's \textls*{\small SSD}s are optimized for reads in the
kilobyte range.

\paragraph{\textls*{\small SSD}s have a low request rate.} The fastest~\cite{fusionio} can
perform on the order of 1 million random reads per second, and many
are limited to less than 100K/s. In contrast, cores may be idle in
Grappa if each node cannot make 100 million random reads per second of
the global heap. Even the individual flash chips are slow, on the order of 40K/s~\cite{micronFlash}.

\paragraph{\textls*{\small SSD}s have high overhead.} Commodity \textls*{\small SSD}s are usually treated
as normal disks: accesses must occur through the kernel, incurring a
high context-switch overhead. Grappa depends on user-mode access to
the network layer. This capability has been explored in recent research~\cite{caulfield:2012}.

@@

We see three ways to address this disparity.
\paragraph{Scaling up the Flash array}
Existing Flash-based \textls*{\small SSD}s are designed for large block requests. An
\textls*{\small SSD} for Grappa could be designed to support high concurrency, small
requests. Where a traditional \textls*{\small SSD} might put multiple Flash chips on
the same bus, or use Flash chips in parallel to build a wider,
high-bandwidth bus, a Grappa \textls*{\small SSD} would try to maximize parallelism by
putting each chip on its own bus. Serial Flash might be used to
minimize pin count requirements, at the cost of some additional
latency. This design could scale to request rates, but the challenge
would be chip count: to support 100M reads per second, an array of
20$\mu$s flash would require 2000 chips. This is possible, but expensive.

\paragraph{Scaling down the processors.}
Alternatively, rather than scaling up the flash array, processors
could be scaled down to match the flash request rate. A low-power
processor with few cores might generate only 1 million references per
second. This could be satisfied with only 20\,x\,20$\mu$s Flash chips.

@@

\paragraph{Buffering requests to increase locality.}
Grappa trades latency for better throughput by aggregating network
communications for better utilization and by migrating synchronization
operations for reduced contention. The same idea could be applied to
increase locality and amortize the cost of Flash accesses. Requests
to common flash pages would be bucketed by flash page address for a
limited time. When a sufficient number of requests have arrived for a page
or when the requests time out, a single page request would be issued to the Flash
layer; its result would be used to satisfy the individual requests.

@@

\paragraph{\textls*{\small PCM} as Main Memory}
Much recent research has explored the use of \textls*{\small PCM} as a main memory
replacement. Grappa would easily be able to take advantage of such a
development. When a Grappa program references low-locality data,
a prefetch request is issued for that data a few cycles before the
thread is context switched in. A \textls*{\small PCM}-based
Grappa would do the same, but would issue the prefetches earlier to
cover the longer latency of \textls*{\small PCM}.

\paragraph{\textls*{\small PCM} as attached storage}

Grappa could also support a \textls*{\small PCM} array in a separate address space (for
instance, on the \textls*{\small PCIe} bus). \textls*{\small PCM} chips available today~\cite{micronPCM} support a
64-bit read in 200 ns; this means that 20 chips acting independently
would be sufficient to reach our target request rate.

@@

\section{Conclusion}
Efficient execution of irregular applications on large datasets is of
growing importance.  Grappa promises to address this need on mass
market clusters.  Addition of non-volatile storage is a natural
extension of such systems.  We ask, what would developing a Pomace for
Grappa -- a Grappa system exploiting non-volatile memory -- entail and
what capability and efficiency improvements might it bring to Grappa's
application space.  Answering this question is an opportunity for
collaboration between our group and the \textls*{\small NVMW} community.

%\subsection{References}
\bibliographystyle{abbrv}
\bibliography{paper}

>>>>

\input{header}
\usepackage{microtype}
\usepackage{placeins}
\begin{document}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{.7in}
\addtolength{\oddsidemargin}{-.25in}
\addtolength{\evensidemargin}{-.25in}
\addtolength{\textwidth}{.5in}
\addtolength{\topsep}{-3pt}
\linespread{.94}

@@

\FloatBarrier
\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.95\columnwidth]{../asplos13/figs/system-overview}
\begin{minipage}{0.95\columnwidth}
  \caption{\label{fig:grappa} Grappa system overview}
\end{minipage}
%\vspace{-6ex}
\end{center}
\end{figure}

@@

\paragraph{\textls*{\small SSD}s assume locality.} In our benchmarks, we find the average request size to be
%\noindent{\bf \textls*{\small SSD}s assume locality.} In our benchmarks, we find the average request size to be
less than 64 bytes. Today's \textls*{\small SSD}s are optimized for reads in the
kilobyte range.

\paragraph{\textls*{\small SSD}s have a low request rate.} The fastest~\cite{fusionio} can
%\noindent{\bf \textls*{\small SSD}s have a low request rate.} The fastest~\cite{fusionio} can
perform on the order of 1 million random reads per second, and many
are limited to less than 100K/s. In contrast, cores may be idle in
Grappa if each node cannot make 100 million random reads per second of
the global heap. Even the individual flash chips are slow, on the order of 40K/s~\cite{micronFlash}.

\paragraph{\textls*{\small SSD}s have high overhead.} Commodity \textls*{\small SSD}s are usually treated
%\noindent{\bf \textls*{\small SSD}s have high overhead.} Commodity \textls*{\small SSD}s are usually treated
as normal disks: accesses must occur through the kernel, incurring a
high context-switch overhead. Grappa depends on user-mode access to
the network layer. This capability has been explored in recent research~\cite{caulfield:2012}.

@@

We see three ways to address this disparity.

\paragraph{Scaling up the Flash array.}
%\noindent{\bf Scaling up the Flash array.}
Existing Flash-based \textls*{\small SSD}s are designed for large block requests. An
\textls*{\small SSD} for Grappa could be designed to support high concurrency, small
requests. Where a traditional \textls*{\small SSD} might put multiple Flash chips on
the same bus, or use Flash chips in parallel to build a wider,
high-bandwidth bus, a Grappa \textls*{\small SSD} would try to maximize parallelism by
putting each chip on its own bus. Serial Flash might be used to
minimize pin count requirements, at the cost of some additional
latency. This design could scale to request rates, but the challenge
would be chip count: to support 100M reads per second, an array of
20$\mu$s flash would require 2000 chips. This is possible, but expensive.

\paragraph{Scaling down the processors.}
%\noindent{\bf Scaling down the processors.}
Alternatively, rather than scaling up the flash array, processors
could be scaled down to match the flash request rate. A low-power
processor with few cores might generate only 1 million references per
second. This could be satisfied with only 20\,x\,20$\mu$s Flash chips.

@@

\paragraph{Buffering requests to increase locality.}
%\noindent{\bf Buffering requests to increase locality.}
Grappa trades latency for better throughput by aggregating network
communications for better utilization and by migrating synchronization
operations for reduced contention. The same idea could be applied to
increase locality and amortize the cost of Flash accesses. Requests
to common flash pages would be bucketed by flash page address for a
limited time. When a sufficient number of requests have arrived for a page
or when the requests time out, a single page request would be issued to the Flash
layer; its result would be used to satisfy the individual requests.

@@

\paragraph{\textls*{\small PCM} as Main Memory.}
%\noindent{\bf \textls*{\small PCM} as Main Memory.}
Much recent research has explored the use of \textls*{\small PCM} as a main memory
replacement. Grappa would easily be able to take advantage of such a
development. When a Grappa program references low-locality data,
a prefetch request is issued for that data a few cycles before the
thread is context switched in. A \textls*{\small PCM}-based
Grappa would do the same, but would issue the prefetches earlier to
cover the longer latency of \textls*{\small PCM}.

\paragraph{\textls*{\small PCM} as attached storage.}
%\noindent{\bf \textls*{\small PCM} as attached storage.}
Grappa could also support a \textls*{\small PCM} array in a separate address space (for
instance, on the \textls*{\small PCIe} bus). \textls*{\small PCM} chips available today~\cite{micronPCM} support a
64-bit read in 200 ns; this means that 20 chips acting independently
would be sufficient to reach our target request rate.

@@

\section{Conclusion}
Efficient execution of irregular applications on large datasets is of
growing importance.  Grappa promises to address this need on mass
market clusters.  Addition of non-volatile storage is a natural
extension of such systems.  We ask, what would developing a Pomace for
Grappa -- a Grappa system exploiting non-volatile memory -- entail and
what capability and efficiency improvements might it bring to Grappa's
application space?  Answering this question is an opportunity for
collaboration between our group and the non-volatile memory research community.

\begin{spacing}{0.8}
\bibliographystyle{abbrv}
\bibliography{paper}
\end{spacing}
